---
title: "HW 4"
author: "Lindsay Payne - UTEID: Lnp832"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document:
    toc: true
---

```{r global_options, echo=FALSE}
knitr::opts_chunk$set(fig.height=3, fig.width=4, warning=FALSE, tidy=TRUE, tidy.opts=list(width.cutoff=60))
```

----------------------


```{r, echo = FALSE, message = FALSE}
# Load in necessary libraries
library(tidyverse)
library(mosaic)
library(boot)
library(knitr)
library(kableExtra)
```

## Problem 1: Iron Bank
```{r, echo = FALSE, fig.width = 5, fig.height = 4}
# Run 100000 Simulations of the test statistic
sim_flagged = do(100000)*nflip(n=2021, prob=0.024)

# Create a graph of the distribution
ggplot(sim_flagged) + 
  geom_histogram(aes(x=nflip), binwidth=1, fill = "darkslategray", color = "black") +
  labs(title = "Probability Distribution of Flagged Trades by Iron Bank",
       x = "Number of Flagged Trades out of 2021",
       y = "Count")
```

```{r, echo = FALSE}
# Calculate the p-value
p_value <- sum(sim_flagged >= 70) / 100000

# Create a table summarizing results
p_value_table <- data.frame(
  Total_Simulations = 100000,
  At_Least_70_Flagged = 190, 
  P_Value = 190 / 100000
)

# Print a table
kable(p_value_table, caption = "Simulation Results for Flagged Trades", align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", "hover"), 
                full_width = FALSE, position = "left")
```

Our null hypothesis is that securities trades from the Iron Bank are flagged at the same 2.4% baseline rate as that of other traders over the long run. Our test statistic is the number of trades by Iron bank employees that were flagged by the SEC's detection algorithm. Higher numbers of flags imply stronger evidence against the null hypothesis, and in our data 70 out of 2,021 trades were flagged. Since only 190 out of the 100,000 Monte Carlo simulations we ran, which assumed the null hypothesis of a 2.4% baseline rate was true, resulted in 70 or more flagged trades, our p-value is p = 0.0019. This value is a decent amount smaller than 0.05, suggesting that the observed number of flagged trades is statistically significant or highly unlikely to have occurred by chance alone, providing strong evidence against our null hypothesis and indicating that Iron Bank employees may be engaging in unusual trading activity.

## Problem 2: Health Inspections
```{r, echo = FALSE, fig.width = 5, fig.height = 4}
# Run 100000 simulations of the test statistic
sim_violation = do(100000)*nflip(n=50, prob=0.03)

# Create a graph of the distribution
ggplot(sim_violation) + 
  geom_histogram(aes(x=nflip), binwidth=1, fill = "darkslategrey", color = "black") +
  labs(title = "Probability Distribution of Health Code Violations",
       x = "Number of Violations in 50 Inspections",
       y = "Count")
```

```{r, echo = FALSE}
# Calculate the p-value
p_value2 <- sum(sim_violation >= 8) / 100000

# Create a table summarizing results
p_value_table2 <- data.frame(
  Total_Simulations = 100000, 
  At_Least_8_Violations = 13, 
  P_Value = 13 / 100000
)

# Print a table
kable(p_value_table2, caption = "Simulation Results for Health Violations", align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", "hover"), 
                full_width = FALSE, position = "left")
```
Our null hypothesis is that restaurants in the city are cited for health code violations at the same 3% baseline rate, on average. Our test statistic is the number of health code violations at Gourmet Bites. Higher numbers of violations imply stronger evidence against the null hypothesis, and in our data 8 out of 50 Gourmet Bites health inspections resulted in a violation. Since only 13 out of the 100,000 simulations we ran, which assumed the null hypothesis violation baseline rate was true, resulted in 8 or more violations, our p-value is p = 0.00013. This value is a decent amount smaller than 0.05, suggesting that the observed number of violations is statistically significant or highly unlikely to have occurred by chance alone, providing strong evidence against the null hypothesis and indicating that Gourmet Bites’ rate of health code violations is significantly higher than the citywide average of 3%.

## Problem 3: Evaluating Jury Selection for Bias
```{r, echo = FALSE}
# Set the expected and observed jury/racial distribution
expected_distribution = c(Group_1 = 0.30, Group_2 = 0.25, Group_3 = 0.20, Group_4 = 0.15, Group_5 = 0.10)
observed_count = c(Group_1 = 85, Group_2 = 56, Group_3 = 59, Group_4 = 27, Group_5 = 13)
```

```{r, echo = FALSE, fig.width = 5, fig.height = 4}
# Use "multinomial sampling" to equal sampling from a named set of categories
num_jurors = 240
simulated_counts = rmultinom(1, num_jurors, expected_distribution)

# Define a function to calculate the chi-squared statistic
chi_squared_statistic = function(observed, expected) {
  sum((observed - expected)^2 / expected)
}

# Calculate chi square statistic for one simulation
chi2_jury = chi_squared_statistic(simulated_counts, num_jurors*expected_distribution)

# Repeat 10000 times
num_simulations = 10000
chi2_sim = do(num_simulations)*{
  simulated_counts = rmultinom(1, num_jurors, expected_distribution)
  this_chi2 = chi_squared_statistic(simulated_counts, num_jurors*expected_distribution)
  c(chi2_jury = this_chi2)
}

# Create a graph of the distribution
ggplot(chi2_sim) + 
  geom_histogram(aes(x=chi2_jury), binwidth=1, fill = "darkslategrey", color = "black") +
  labs(title = "Probability Distribution of Chi-Square Statistics",
       x = "Chi-Square Statistic",
       y = "Count")

# Calculate the observed chi squared statistic
my_chi2 = chi_squared_statistic(observed_count, num_jurors*expected_distribution)

# Calculate the p-value of the test statistic
p_value3 <- sum(chi2_sim >= 12.42639) / 10000
```

Our null hypothesis, H0, is that the distribution of jurors empaneled by this judge matches the county's population proportions, meaning the jury selection process is fair and random. Our test statistic, 𝑇 , is the chi-square statistic or how much the observed jury selection differs from the expected population proportions. A higher difference implies stronger evidence against the null hypothesis, and in our data this chi-square statistic was 12.42639. Since 139 out of the 10,000 simulations we ran, which assumed the null hypothesis of a fair jury selection was true, resulted in a chi-square statistic of 12.42639 or more, our P(T|H0) or p-value is p = 0.0153. This value is less than 0.05, suggesting that the observed jury distribution is significantly different from the county’s population proportions. 

However, this does not necessarily entail systematic bias in jury selection because many other factors or explanations may exist. For example, differences in courtroom attendance and response rates. Some racial groups may be less likely to respond to jury summons or appear in court which could be further investigated by looking at response rates across different racial/ethnic groups. Additionally some groups may be more likely to qualify for exemptions, like those with financial hardships, which could be investigated by comparing rates of exemptions across groups.

## Problem 4: LLM Watermarking
```{r, echo = FALSE, fig.width = 5, fig.height = 4}
# Step 1: Read in sentences
brown_sentences <- readLines("brown_sentences.txt")

# Step 2: Function to preprocess text and count letter occurrences per sentence
process_sentence <- function(text) {
  clean_text <- gsub("[^A-Za-z]", "", text)
  clean_text <- toupper(clean_text)
  letter_counts <- setNames(rep(0, 26), LETTERS)
  for (letter in strsplit(clean_text, "")[[1]]) {
    letter_counts[letter] <- letter_counts[letter] + 1
  }
  return(letter_counts)
}

# Step 3: Apply preprocessing and letter counting
letter_counts <- sapply(brown_sentences, process_sentence, simplify = FALSE)

# Read in letter frequencies and convert to named vector
letter_freq <- read.csv("letter_frequencies.csv")
letter_freq <- setNames(letter_freq$Probability, letter_freq$Letter)

# Step 4: Function to calculate expected letter counts for sentence length
calculate_expected_counts <- function(sentence_length, letter_freq) {
  expected_counts <- setNames(rep(0, 26), LETTERS)
  expected_counts[names(letter_freq)] <- sentence_length * letter_freq
  return(expected_counts)
}

# Step 5: Function to calculate chi-squared statistic
calc_chi_squared <- function(observed, expected) {
  sum((observed - expected)^2 / expected)
}

# Calculate chi-squared statistics for each sentence
chi_squared_stats <- sapply(letter_counts, function(observed_counts) {
  sentence_length <- sum(observed_counts)  
  expected_counts <- calculate_expected_counts(sentence_length, letter_freq)
  
  # Make sure observed counts match expected names
  observed_counts <- observed_counts[names(expected_counts)]
  calc_chi_squared(observed_counts, expected_counts)
})

# Create a data frame to help with visualization
chi_squared_df <- data.frame(chi_squared = chi_squared_stats)

# Step 6: Create a graph of the distribution
ggplot(chi_squared_df) + 
  geom_histogram(aes(x=chi_squared), bins = 30, fill = "darkslategrey", color = "black") +
  labs(title = "Probability Distribution of Chi-Square Statistics",
       x = "Chi-Square Statistic",
       y = "Count")
```

```{r, echo = FALSE}
# Convert 10 sentences into an R vector
test_sentences <- c(
  "She opened the book and started to read the first chapter, eagerly anticipating what might come next.",
  "Despite the heavy rain, they decided to go for a long walk in the park, crossing the main avenue by the fountain in the center.",
  "The museum's new exhibit features ancient artifacts from various civilizations around the world.",
  "He carefully examined the document, looking for any clues that might help solve the mystery.",
  "The students gathered in the auditorium to listen to the guest speaker's inspiring lecture.",
  "Feeling vexed after an arduous and zany day at work, she hoped for a peaceful and quiet evening at home, cozying up after a quick dinner with some TV, or maybe a book on her upcoming visit to Auckland.",
  "The chef demonstrated how to prepare a delicious meal using only locally sourced ingredients, focusing mainly on some excellent dinner recipes from Spain.",
  "They watched the sunset from the hilltop, marveling at the beautiful array of colors in the sky.",
  "The committee reviewed the proposal and provided many points of useful feedback to improve the project's effectiveness.",
  "Despite the challenges faced during the project, the team worked tirelessly to ensure its successful completion, resulting in a product that exceeded everyone's expectations."
)
```

```{r, echo = FALSE}
# Calculate chi-squared statistics for test sentences by reusuing Part A's function
test_chi_squared_stats <- sapply(test_sentences, function(sentence) {
  observed_counts <- process_sentence(sentence) 
  sentence_length <- sum(observed_counts)
  expected_counts <- calculate_expected_counts(sentence_length, letter_freq)
  observed_counts <- observed_counts[names(expected_counts)]  
  calc_chi_squared(observed_counts, expected_counts)  
})

# Calculate p-values by comparing with Part A's chi-squared distribution. Use proportion of normal sentences with chi-squared ≥ observed
p_values <- sapply(test_chi_squared_stats, function(stat) {
  mean(chi_squared_stats >= stat)
})

# Create a table of p-values
p_value_table4 <- data.frame(Sentence = 1:10, P_Value = round(p_values, 3))

# Print the table
kable(p_value_table4, caption = "P-Values for Each Sentence", align = "c") %>%
  kable_styling(bootstrap_options = c("striped", "condensed", "hover"), 
                full_width = FALSE, position = "left")
```

The sentence that was produced by an LLM is sentence 6 because it has the smallest p-value, which is 0.009. This low p-value suggests that the sentence’s letter frequency distribution differs significantly from typical English text, signifying the potential presence of an LLM watermark. Additionally, this is the only p-value calculated that is smaller than 0.05, making it statistically significant and unlikely under the null hypothesis that the sentences follow normal English letter frequency.
